Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2023-07-17 ba_resubmit/response_to_reviewers2.tex

Thank you for the thoughtful feedback. We have made several broad changes to address the Associate Editor's overall critiques, and have addressed each specific comment from the reviewer.

\section{AE Comments}

\pointRaised{AE}{I must add that reading the paper still feels like a chore, in part because the notation is so heavy, which may be hard to avoid, but also, in part, because some wording choices make certain sentences hard to parse, which can certainly be improved.}

\reply{We have made several efforts to simplify notation in this revision.

\begin{itemize}
	\item Instead of referring to files as $\mathbf{X}_1$ and $\mathbf{X}_2$, we are now using $A$ and $B$. This particularly simplifies the notation in Section 4.2.

	\item At the end of Section 4.1, we have rewritten the likelihood function in terms of the agreement patterns and the summary statistics.
	This likelihood is then used throughout Section 4.2 in the discussion of efficient posterior inference. We hope this expression clarifies the meaning of "preservation of weight (Comment 5), and strengthens the argument for $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N} \}$ being summary statistics for $\bm{m}$, $\bm{u}$, and $\bm{Z}$ (Comment 4).

	\item To denote when the $(i,j)$ record pair exhibits agreement pattern $p$, we have removed the $(i, j) \in h_p$ notation for $\gamma_{ij} = h_p$.

@@

	\pointRaised{R}{%
	For the missing data treatment now included just before Section 2.1, is the assumption truly missing at random, or missing completely at random? Or, does this distinction not matter because of the independence assumed across elements of the comparison vector?}

	\reply{% 
		The assumption is missing completely at random. You are correct that this is a result of the independence assumed across elements of the comparison vector. This has been clarified.
	}

	\pointRaised{R}{%
	Appendix 8.2: I appreciate the streamlining of the discussion of the proposed algorithms and think it was a good choice to move the full derivation to an appendix. However, I do not follow the re-expression of the pmf for $\Gamma_{.j}$. First, the opening square brackets are still misplaced
	throughout. In the second line of this derivation, the authors divide the expression by a product of the element-wise conditional probabilities of a match in the comparison vector (u), raised by an indicator that the element of the comparison vector equals a particular value. Perhaps I misunderstand, but I believe this product is not constant in $\Gamma_{.j}$ or $u$, and so the total expression is not proportional to the line above, as interpreted as a pmf. I think that the end effect is that the last line of the newly expressed pmf is missing a factor of $\prod_i \prod_f \prod_l u_{fl}^{I(\gamma_{ij}^f = l) I_{obs}(\gamma_{ij}^f)}$ regardless of the value of $z_j$.}

\reply{%
	The square brackets have been corrected, and we apologize for the oversight. We have clarified derivation of the likelihood for $Z_j$, separately handling the case when $B_j$ has a match and when it does not.
}

@@

	\pointRaised{R}{%
Appendix 8.2: Thank you for including the details on integrating out pi from the full conditionals. They surprised me. I had assumed that the authors had integrated out pi in the prior for Z, as this is what Sadinle (2017) had done to form the “beta prior for bipartite matchings”. Does the alternate approach presented in this paper provide a different algorithm than directly integrating the prior distribution? Is the presented alternative approach justified?}

\reply{%
	We realize that the integration of pi out of the full conditionals is not necessary for our sampler. For clarity, we decided to redo the simulations using a Gibbs sampler that samples $\pi$ from its full conditional, and then samples $Z$ from its full conditionals. We obtained equivalent results, which is unsurprising, because with high numbers of observations, the posterior of the beta distribution is highly concentrated around its mean.

	Therefore, we have updated the paper with full conditionals for the standard Gibbs sampler. We leave the remark about marginalization in the Appendix to facilitate comparisons to Sadinle's method.

@@

}

\pointRaised{R}{%
	I do not understand the statement just below Equation (10): “When j has no match in $X_1$, we write $(n_1 +j, j) \in h_{P+1}$” My understanding of these patterns is that they are based on observed comparison vectors without consideration of Z (matches). In the second paragraph of Section 4.3, the H notation includes the matches (Z), in notation and definition that seems to conflict with the statement just below Equation (10). This also comes into play in Equation (16)}

	\reply{%
	You are correct that the comparison vectors are created (and therefore, patterns are assigned) without regard to $Z$. This notation was created purely to be able to denote when a record was left unlinked during the Gibbs Sampler as in Equation 16 in the submitted draft.

	I can see why this is confusing. Therefore, we have removed that line, and changed the notation for record $B_j$ being unmatched to just be''otherwise".
	}

@@

\pointRaised{R}{%
	Third paragraph of Section 4.1: The authors claim they are computing “sufficient statistics”. What exactly are these statistics sufficient for?
	}

	\reply{%
		We show in the revised Section 4.2 that we can write conditional likelihoods  and posterior updates for $\bm{m}$, $\bm{u}$, and $\bm{Z}$ using the statistics in $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N} \}$. However, we cannot express the full likelihood in (14) through these statistics. Therefore, we have revised to text to say ``summary statistics" instead of  ``sufficient statistics."}

@@

\pointRaised{R}{%
	First paragraph of Section 4.2: The authors state: “Posterior calculations still attribute the appropriate weight to all records through the summary statistics…” What is meant by the term “weight”? Which records are appropriately weighted – those in X2?}

\reply{%

	We have switched the order of Sections 4.2 and 4.3, so that we first discuss hashing, then posterior inference, then chunkwise computation of the comparison matrix, and then SEI. With this new ordering, and with the newly provided equations (15 -19) it is clearer that the contribution of \emph{all} record pairs is recorded through the summary statistics in $\mathcal{N}$.
}

@@

\pointRaised{R}{%
	Where does $R^{SEI, cd}$ come into play in the partitioned algorithm presented in Equations (13) and (14)? I recommend that the authors either refrain from suppressing the SEI notation or further explain how the SEI algorithm has changed the quantities in these equations.}

\reply{%
	We have reordered Section 4 so that all posterior inference is presented before SEI. This makes it clear that the posterior updates in (16a), (16b), and (18) depend only on $\mathcal{N}$. SEI only affects the step shown in (19).
}

\pointRaised{R}{%
	Section 4.2: I appreciate the practical advice about choosing S for the SEI method. However, this choice seems arbitrary in the absence of further discussion/evidence. Given that the primary novelty of the manuscript is in methods to speed and otherwise improve computation, I am surprised that this aspect of computational innovation is presented with virtually no theoretical or empirical exploration. Presumably the SEI method has some sort of accuracy trade-off, as the authors warn that linkage results may be “distorted” if S is low. However, this trade-off is not quantified or even discussed in practical terms beyond the terse recommendation to choose S=10.
}

\reply{% 
	We have provided an additional simulation study in Section 5.3 to show the sensitivity to different choices of $S$.
}
\pointRaised{R}{%
	Page 5, 2 sentences before equation (4): I believe the sum should be of $I(Z_j \leq n_1)$, not $I(Z_j \leq n_1 +1)$.}

\reply{%
	This has been corrected, thank you.
}

### 2023-07-17 ba_resubmit/kundinger_resubmit2.tex
beka and jerry edits

\begin{abstract}
%Recently, researchers have developed Bayesian versions of the Fellegi-Sunter model for record linkage. These have the crucial advantage of quantifying uncertainty from imperfect linkages.  
Within the field of record linkage, Bayesian methods have the crucial advantage of quantifying uncertainty from imperfect linkages. However, current implementations of Bayesian Fellegi-Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks. To address these computational considerations, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}. Specifically, we use independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead. This also allows us to complete pairwise record comparisons over large data files through parallel computing and to reduce memory costs through a new technique called storage efficient indexing. Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
%
%\textcolor{blue}{We propose a variation on Bayesian Fellegi-Sunter models that we call fast beta linkage, or \texttt{fabl}, and is motivated by computational considerations.  Our method extends that of \cite{sadinle_bayesian_2017}, who extended the early work of \cite{fellegi_theory_1969}.}

@@

%% ** Keywords **
\begin{keyword}%[class=MSC]
\kwd{Bayesian methods}
\kwd{distributed computing}
\kwd{entity resolution}
\kwd{hashing}
\kwd{record linkage}
\end{keyword}

@@

Many statistical record linkage methods are extensions of the seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}. Specifically, \cite{fellegi_theory_1969} created comparison vectors for each pair of records in the data files and independently classified each pair as a match or a non-match using a likelihood ratio test. Recent work in the statistical literature has extended this approach for a wide variety of applications \citep{Winkler1990, fair2004generalized, wagner2014person, gill2003english, enamorado2019using, aleshinguendel2021multifile}. Additionally, some methods directly model the linkage variables \citep{steorts_bayesian_2016, marchant_distributed_2019, betancourt2021prior}, but in this paper, we build on the contributions to the comparison vector approach.

@@

The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular mainly for its mathematical simplicity, but can be unreasonable in practice. In many situations, we know that there are no duplications within a data file, meaning that each record in one file should be linked with at most one other record in the other file. Thus, when the procedure results in many-to-one matches, some of these links must be false. Many extensions to \cite{fellegi_theory_1969} resolve these false matches as a post-processing step \citep{jaro1989}, but this model misspecification can still lead to poor results \citep{sadinle_bayesian_2017}.

Alternatively, one can embed one-to-one matching requirements into the model specification itself \citep{gutman2013bayesian, liseo_2011}, at an additional computational cost. \cite{Larsen05} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space. \cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method is computationally intensive and, to our knowledge, has not been applied on data files with more than 100 records. \cite{sadinle_bayesian_2017} proposed the Beta Record Linkage model (\texttt{BRL}), using a prior distribution over the space of bipartite matchings to strictly enforce one-to-one requirements throughout a Gibbs sampler. Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeler can weigh the relative importance of false positives and false negatives, and identify records pairings to be decided through clerical review. \texttt{BRL} was shown to work on larger tasks than previous one-to-one methods, but in our experience, it becomes slow when applied to files with more than a few thousand records.

In this paper, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. Following the suggestion in \cite{wortman2019}, we relax the one-to-one matching requirement of \texttt{BRL} and use independent priors over the matching space. This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing. These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with minimal loss of accuracy. In particular, computation time under \texttt{BRL} grows quadratically, with the size of each data file, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller data file.

%\textcolor{blue}{Moreever, we propose storage efficient indexing (SEI), which reduces memory costs.}
%\textcolor{blue}{More specifically, we use both hashing and SEI to increase the scale of the linkage tasks without the use of blocking. This can be useful when there is no reliable blocking field available or one desires estimates of model parameters for the entire sample in question. In practice, \texttt{fabl} can be combined with blocking, but all derivations, simulations, and case studies are presented here without blocking. We illustrate our methodology on two case studies and two simulation scenarios, where Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.}

%We demonstrate \texttt{fabl} provides accurate estimation of the linkage structure and greatly enhanced speed. Open source software to use \texttt{fabl} in \texttt{R} is available through \href{https://github.com/briankundinger/parlrdev}{Github}.

@@

\section{Review of Prior Work}
\label{sec:review-of_prior-work}

%Consider two data files $A$ and $B$ containing records $\{x_{1i}\}_{i=1}^{n_A}$ and $\{x_{2j}\}_{j=1}^{n_B}$ respectively.

Consider two data files $A$ and $B$, consisting of records $\{A_i: i = 1, \ldots, n_A\}$ and $\{B_j: j = 1, \ldots, n_B\}$. Suppose the files have $F$ fields in common to be used for linkage, and that these fields take on values $A_{if}$ and $B_{jf}$ respectively. Without loss of generality, denote files such that $n_A \geq n_B$. Under bipartite matching, the set of matches across data files can be represented in two equivalent ways. First, we may use a matrix $\Delta \in \{0, 1\}^{n_A \times n_B}$, where
\begin{align}
	\Delta_{ij} =
	\begin{cases}
		1, & \text{if records}\;  A_i \; \text{and}\; B_j \; \text{refer to the same entity}; \\
		0, & \text{otherwise}.\\
	\end{cases}
\end{align}
This sparse matrix representation can become cumbersome for large linkage tasks. More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_B})$ for the records in $B$ such that 
\begin{align}
	Z_{j} =
	\begin{cases}
		i, & \text{if records}\;  A_i \; \text{and}\; B_j  \; \text{refer to the same entity}; \\
		n_A + j, & \text{if record}\;  B_j \; \text{does not have a match in}\; A. \\
	\end{cases}
\end{align}
We can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise.

{\color{red}{The following paragraph is pretty unnecessary, I never use this notation}}
Denote the set of matches by $\bm{M} = \{(i,j): i \leq n_A, j \leq n_B, \Delta_{ij} = 1\}$, and the set of non-matches by 
$\bm{U} =  \{(i,j): i \leq n_A, j \leq n_B, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'' and to record pairs that are estimated as non-matches as ``non-links.''

%\subsection{Comparison Vectors}
%\label{comparison-vectors}

{\color{red}{Beka suggested eliminating the following two paragraphs. Thoughts? I think the first one is good, the second seems unneccesary.}}

Intuitively, matching records (those that refer to the same entity) should be similar; records that are non-matching should be dissimilar. \cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\gamma_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$ Denote the number of criteria for comparing records by $F$, such that $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^F).$ {\color{red}{We use $\Gamma_{ij}$ when referring to the vector as a random variable, and $\gamma_{ij}$ when referring to observed data. We collect the vectors $\gamma_{ij}$ as the comparison matrix $\gamma \in \mathbb{R}^{n_A n_B \times F}$}}.

%Frequently, $\gamma_{ij}$ consists of one comparison for each feature shared between the two data files.

The simplest way to compare any particular feature for two records is to check for exact agreement, and this is commonly used for categorical features. For example, if zip code is linking feature $f$, we can set $\gamma_{ij}^f=1$ when the zip codes for records $A_i$ and $B_j$ agree exactly, and set $\gamma_{ij}^f=2$ when they do not. For numerical features, we can use absolute difference between the two feature values. For example, if age is linking field $f$, we can set $\gamma_{ij}^f = 1$ when the ages for records $A_i$ and $B_j$ match exactly, $\gamma_{ij}^f = 2$ when the ages for records $A_i$ and $B_j$ are within one year but not equal, and $\gamma_{ij}^f = 3$ when the ages are two or more years apart. For text features, such as names, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}. We then set thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.

More generally, let $\mathcal{S}_f(i,j)$ denote a similarity measure for feature $f$ of records $A_i$ and $B_j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f$ intervals denoted by $I_{f1}, \ldots, I_{fL_f}$. Here, $I_{f1}$ represents the highest level of agreement (including complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement). Thus, we can construct comparison vectors such that $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choices of $I_{fl}$ are application specific, as we discuss in the simulation and case studies.

{\color{red}{
In the construction of comparison vectors, it is common to encounter missing information in record $A_i$ or $B_j$. As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values. We assume that this missingness occurs completely at random (MCAR, per \cite{LittleRubin2002}). To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise. With the MCAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data. }
}

%For more complex measurements, we can take into account partial agreement to more richly characterize the comparison;

%
%In this paper, we use hashing and a new technique called storage efficient indexing to increase the scale of the linkage tasks we can undertake without blocking. This is useful when there is no reliable blocking field available, or one desires estimates of model parameters for the entire sample in question. In practice, \texttt{fabl} can be combined with blocking, but all derivations, simulations, and case studies are presented here without blocking.

\subsection{Fellegi-Sunter Models}
\label{fellegi-sunter}

The seminal \cite{fellegi_theory_1969} model employs two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent. Using these assumptions, \cite{winkler_state_1999}, \cite{jaro1989}, \cite{larsen_2001}, \cite{enamorado2019using} and others model the comparison data through mixture models of the form
\begin{subequations}
\begin{align}
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \label{eqn:fs_model} \\
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \label{eqn:m_dist}\\
	&\Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(\lambda). \label{eqn:u_dist}
\end{align}
\end{subequations}
Here, $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and non-matching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $\lambda$ is the marginal probability that a record pair is a match. When using comparison vectors with discrete agreement levels, $\mathcal{M}$ and $\mathcal{U}$ are collections of independent multinomial distributions for each linkage feature. Accordingly, $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$. The $\bm{u}$ parameters are defined similarly, with $u_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 0)$.

Within this framework, record pairs are independently classified as matches and non-matches based on the estimated parameters. However, such independent classifications often leads to links that violate one-on-one matching assumptions, requiring post-processing to achieve desirable results. To address this issue, \cite{jaro1989} proposed an optimization technique using the estimated model parameters to produce a bipartite matching. \cite{sadinle_bayesian_2017} later showed that this method is equivalent to a maximum likelihood estimate.

%This prompted \cite{jaro1989} to formulate an optimization problem that produces a bipartite matching maximizing the sum of the Fellegi-Sunter weights among matched pairs.

%These unknown parameters are often estimated using the EM algorithm \cite{winkler_state_1999}.

@@

In contrast to the prior over the vector $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$. However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use the fast Beta prior as follows. For each $Z_j$, we have
\begin{align}
	\label{eqn:fast_beta_prior}
	p(Z_j = q| \pi) &= \begin{cases} 
	\frac{1}{n_A}\pi,  & q \leq n_A; \\
	1-\pi, &  q  = n_A + j; \\
\end{cases} \\
\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}) \notag.
\end{align}
We can interpret (\ref{eqn:fast_beta_prior}) as follows: record $B_j$ has some match in $A$ with probability $\pi$, and each record $A_i$ is equally likely to be that match. The hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about  the proportion of records in $B$ that have matches in $A.$

>>>>

\begin{abstract}
%Recently, researchers have developed Bayesian versions of the Fellegi-Sunter model for record linkage. These have the crucial advantage of quantifying uncertainty from imperfect linkages.  
Within the field of record linkage, Bayesian methods have the crucial advantage of quantifying uncertainty from imperfect linkages. However, current implementations of Bayesian Fellegi-Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks. To address these computational considerations, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}. Specifically, we use independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead. This also allows us to complete pairwise record comparisons over large data files through parallel computing and to reduce memory costs through a new technique called storage efficient indexing. Through simulations and two case studies, we show that \texttt{fabl} can have markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
%
%\textcolor{blue}{We propose a variation on Bayesian Fellegi-Sunter models that we call fast beta linkage, or \texttt{fabl}, and is motivated by computational considerations.  Our method extends that of \cite{sadinle_bayesian_2017}, who extended the early work of \cite{fellegi_theory_1969}.}

@@

%% ** Keywords **
\begin{keyword}%[class=MSC]
\kwd{data fusion}
\kwd{distributed computing}
\kwd{entity resolution}
\kwd{hashing}
\kwd{record linkage}
\end{keyword}

@@

Many probabilistic record linkage methods are rely on the seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}. In their approach, the data analyst first creates comparison vectors for each pair of records in the data files. These vectors indicate how similar the records are on a set of variables measured in both files, known as the linkage variables.  Using these comparison vectors, the analyst classifies each pair as a match or nonmatch using a likelihood ratio test. Crucially, these decisions are made independently for each pair. The \cite{fellegi_theory_1969} approach has been extended for a wide variety of applications \citep[e.g.,][]{Winkler1990, fair2004generalized, wagner2014person, gill2003english, enamorado2019using, aleshinguendel2021multifile}. An alternative paradigm is to model the linkage variables directly \citep[e.g.,][]{tancredi2011hierarchical, steorts_bayesian_2016, marchant_distributed_2019, betancourt2021prior}. In this article, we build on the contributions to the comparison vector approach.

@@

The independent pairwise matching assumption from \cite{fellegi_theory_1969} is popular mainly for its mathematical simplicity. However, In many situations, there are no duplications within a data file, meaning that each record in one file should be linked with at most one other record in the other file. Thus, when the procedure results in many-to-one matches, some of these links must be false. To address this potential problem, data analysts typically use an additional post-processing step, for example, applying an optimization routine that turns the list of linked records into a bipartite matching \citep{jaro1989}, but this model misspecification can still lead to poor results \citep{sadinle_bayesian_2017}.

Alternatively, analysts can embed one-to-one matching requirements into the model specification, at an additional computational cost. \cite{Larsen05} employed a Metropolis-Hastings algorithm to only allow sampling matches that respected one-to-one assumptions, but such algorithms exhibit slow mixing due to the combinatorial nature of the constrained matching space. \cite{fortunato_2010} used simulated annealing to target the space of matches permitted under the one-to-one constraint, but the method is computationally intensive and, to our knowledge, has not been applied on data files with more than 100 records. \cite{sadinle_bayesian_2017} proposed the Beta Record Linkage model (\texttt{BRL}), using a prior distribution over the space of bipartite matchings to strictly enforce one-to-one requirements throughout a Gibbs sampler. Additionally, he introduced a class of loss functions that allows for a flexible estimation of the linkage structure, such that the modeler can weigh the relative importance of false positives and false negatives, and identify records pairings to be decided through clerical review. \texttt{BRL} has been shown to work on larger tasks than previous one-to-one methods, but in our experience, it becomes slow when applied to files with more than a few thousand records.

In this article, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. {\color{red}{We use independent priors for the matching status of each record, inducing a dependency that is stronger than the record pair independence in the original \cite{fellegi_theory_1969} but weaker that the one-to-one requirement of \cite{sadinle_bayesian_2017}.}} This allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing. These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with similar levels of accuracy. In particular, computation time under \texttt{BRL} grows quadratically, with the size of each data file, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller data file.

%Following the suggestion in \cite{wortman2019}, we relax the one-to-one matching requirement of \texttt{BRL} and use independent priors over the matching space.

@@

\section{Review of Fellegi-Sunter Approaches for Record Linkage}
\label{sec:review-of_prior-work}

Consider two data files $A$ and $B$ comprising $n_A$ and $n_B$ records, respectively, and including $F$ linkage variables measured in both files. For $i=1, \dots, n_A$, let record $i$ be given by $A_i = (A_{i1}, \dots, A_{iF})$, so that $A = (A_i : i = 1, \dots, n_A)$.  Similarly, for $j=1, \dots, n_B$, let record $j$ be given by $B_j = (B_{j1}, \dots, B_{jF})$, so that $B = (B_j : j = 1, \dots, n_B)$.  Without loss of generality, denote files such that $n_A \geq n_B$.

%Consider two data files $A$ and $B$, consisting of records $\{A_i: i = 1, \ldots, n_A\}$ and $\{B_j: j = 1, \ldots, n_B\}$. Suppose the files have $F$ fields in common to be used for linkage, and that these fields take on values $A_{if}$ and $B_{jf}$ respectively.

Intuitively, matching records (those that refer to the same entity) should have similar values of the linking variables; records that are nonmatching should have dissimilar values. \cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\gamma_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$  Specifically, they define $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^F),$ where each $\gamma_{ij}^f$ is a value indicating the similarity of field $f$ for records $A_i$ and $B_j$. For notational convenience, we define the comparison matrix $\gamma \in \mathbb{R}^{n_A n_B F}$ as the collection of all record pairs' $\gamma_{ij}$.

When linking variable $f$ is categorical, a common way to define $\gamma_{ij}^f$ is an indicator for exact agreement. For example, if zip code is linking variable $f$, we can set $\gamma_{ij}^f=1$ when the zip codes for records $A_i$ and $B_j$ agree exactly, and set $\gamma_{ij}^f=2$ when they do not. For numerical linking variables, we can use the absolute difference of the two values. For example, if age is linking variable $f$, we can set $\gamma_{ij}^f = 1$ when the ages for records $A_i$ and $B_j$ match exactly, $\gamma_{ij}^f = 2$ when the ages are within one year but not equal, and $\gamma_{ij}^f = 3$ when the ages are two or more years apart. For text variables like names, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}. We then set thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.

More generally, let $\mathcal{S}_f(i,j)$ denote a similarity measure for linking variable $f$ of records $A_i$ and $B_j.$ The range of $\mathcal{S}_f$ can be divided into $L_f$ intervals denoted by $I_{f1}, \ldots, I_{fL_f}$. Here, $I_{f1}$ represents the highest level of agreement (including complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement). Thus, we can construct comparison vectors such that $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choices of $I_{fl}$ are application specific, as we discuss in the simulation and case studies.

In the construction of comparison vectors, it is common to encounter missing information in record $A_i$ or $B_j$. As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values. We assume that this missingness occurs completely at random (MCAR, per \cite{LittleRubin2002}). To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise. With the MCAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data.

%\subsection{Fellegi-Sunter Models}
%\label{fellegi-sunter}

Having defined comparison vectors, we now turn to the \cite{fellegi_theory_1969} model for record linkage.  We begin with notation for defining linkages. Under bipartite matching, the set of matches across $A$ and $B$ can be represented in two equivalent ways. First, we may use a matrix $\Delta \in \{0, 1\}^{n_An_B}$, where
\begin{align}
\Delta_{ij} =
\begin{cases}
1, & \text{if records}\;  A_i \; \text{and}\; B_j \; \text{refer to the same entity}; \\
0, & \text{otherwise}.\\
\end{cases}
\end{align}
This sparse matrix representation can become cumbersome for large linkage tasks. More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_B})$ for the records in $B$ such that
\begin{align}
Z_{j} =
\begin{cases}
i, & \text{if records}\;  A_i \; \text{and}\; B_j  \; \text{refer to the same entity}; \\
n_A + j, & \text{if record}\;  B_j \; \text{does not have a match in}\; A. \\
\end{cases}
\end{align}
We can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise. When presenting the models, we use both representations for convenience.

Let $\Gamma_{ij}$ represent a random variable for the comparison vector for arbitrary $A_i$ and $B_j$.  Thus, $\gamma_{ij}$ is a realization of $\Gamma_{ij}$.  For modeling the collection of $n_An_B$ random variables $\Gamma_{ij}$, \cite{fellegi_theory_1969} employ two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent. Using these independence assumptions,  one specifies a mixture model for each $\Gamma_{ij}$ \citep[e.g., as in ][]{winkler_state_1999,jaro1989,larsen_2001,enamorado2019using}.  We have
\begin{subequations}
\begin{align}
&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \label{eqn:fs_model} \\
&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \label{eqn:m_dist}\\
&\Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(\lambda). \label{eqn:lambda_dist}
\end{align}
\end{subequations}
Here, $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and nonmatching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $\lambda$ is the marginal probability that a record pair is a match. When using comparison vectors with discrete agreement levels, $\mathcal{M}$ and $\mathcal{U}$ are collections of independent multinomial distributions for each linkage feature. Accordingly, $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = p(\gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$. The $\bm{u}$ parameters are defined similarly, with $u_{fl} = p(\gamma_{ij}^f = l|\Delta_{ij} = 0)$.

%Within this framework, record pairs are independently classified as matches and nonmatches based on the estimated parameters. However, such independent classifications often leads to links that violate one-on-one matching assumptions, requiring post-processing to achieve desirable results. To address this issue, \cite{jaro1989} proposed an optimization technique using the estimated model parameters to produce a bipartite matching. \cite{sadinle_bayesian_2017} later showed that this method is equivalent to a maximum likelihood estimate.

@@

Instead of the prior over $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$. However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use the fast Beta prior as follows. For each $Z_j$, we have
\begin{subequations}
\begin{align}
	p(Z_j = q| \pi) &= \begin{cases} 
	\frac{1}{n_A}\pi,  & q \leq n_A;\\
	1-\pi, &  q  = n_A + j;  \\
\end{cases} \label{eqn:fast_beta_prior}  \\
\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}) \label{eqn:fast_beta_prior2}.
\end{align}
\end{subequations}
We can interpret (\ref{eqn:fast_beta_prior}) as follows: record $B_j$ has some match in $A$ with probability $\pi$, and each record $A_i$ is equally likely to be that match. The hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about  the proportion of records in $B$ that have matches in $A.$

In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_A, n_A +j\}$ is a priori equally likely for $Z_j$. This amounts to a prior probability of $n_A / (n_A + 1)$ that record $B_j$ has a match in $A$. In our preliminary studies, the this highly informative prior weighting on matching can result in overly optimistic match rates. Hence, we prefer (\ref{eqn:fast_beta_prior}). We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_A \right)$ random variable.
